---
title: "Larsen&Shirey_ELE2021_IndependentAnalysis"
author: "Michael W Belitz"
date: "3/24/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Reanalysis of Larsen and Shirey 2021 Ecology Letters
Original technical comment of interest: https://doi.org/10.1111/ele.13731

In a response to the Larsen and Shirey (2021) technical comment, Fric et al. (2021) make the claim that "... we replicated their [LS] procedures (Supplement S5), finding that LS confused their Onset and Termination results."

Below is a independent reanalysis of Larsen and Shirey's (2021) analysis to see if LS did indeed confuse their Onset and Termination results. We use the same estimated phenology metrics of as Larsen and Shirey (2021), but generate our own independent linear model results for onset and termination coefficients. 

The steps and code for this reanalysis are demonstrated below.

## Read in data
```{r data, echo=TRUE}
# Get data from Larsen & Shirey's GitHub
githubURL <- "https://github.com/RiesLabGU/Larsen-Shirey2020_EcoLettersComment/blob/master/data/phenometrics.RData?raw=TRUE"
load(url(githubURL))
```

## Examine data
Note that the phenology data has later values for termination than for onset. This makes biological sense and confirms that at the stage of estimating phenology metrics, there was no error. 
```{r examinedata, echo=TRUE}
head(pheno.data)
```

## Write functions to generate species-specific Linear models
For each of the 22 species included in the Larsen and Shirey re-analysis, fit a linear model with latitude and year as the predictor variables. Onset and Termination are the response variables. Model coeficcients and p-values of the latitude term are extracted. 

```{r functions, echo=TRUE}
# Onset
onset_coef_fun <- function(spp_name){
  
  mdf <- filter(pheno.data, name == spp_name) # filter to spp.
  spp_lm <- lm(onset~rndLat+year, data = mdf) # fit linear model
  
  out_df <- data.frame(
    name = spp_name,
    onset_coef = spp_lm$coefficients[['rndLat']], # extract latitude coef
    p_value = summary(spp_lm)$coefficients[,4][['rndLat']] # extract latitude p-value
    )
  
  return(out_df)
}

# Termination
offset_coef_fun <- function(spp_name){
  
  mdf <- filter(pheno.data, name == spp_name)
  spp_lm <- lm(term~rndLat+year, data = mdf)
  
  out_df <- data.frame(
    name = spp_name,
    offset_coef = spp_lm$coefficients[['rndLat']],
    p_value = summary(spp_lm)$coefficients[,4][['rndLat']]
  )
  
  return(out_df)
}

```

## Run function on the 22 species included in reanalysis.
```{r runFunctions, echo=TRUE}

# get a list of the 22 spp
spp_list <- unique(pheno.data$name)

### get model coef for each spp

#Onset First
onset_coef_list <- lapply(spp_list, onset_coef_fun)
onset_coef_df <- bind_rows(onset_coef_list) %>% 
  mutate(slope = case_when(onset_coef > 0 & p_value < 0.05 ~ "positive",
                           onset_coef < 0 & p_value < 0.05 ~ "negative",
                           p_value >= 0.05 ~ "not-sig"))

onset_coef_sum <- onset_coef_df %>% 
  group_by(slope) %>% 
  summarise(slope_count = n()) %>% 
  mutate(phenometric = "Onset")


#Offset
offset_coef_list <- lapply(spp_list, offset_coef_fun)
offset_coef_df <- bind_rows(offset_coef_list) %>% 
  mutate(slope = case_when(offset_coef > 0 & p_value < 0.05 ~ "positive",
                           offset_coef < 0 & p_value < 0.05 ~ "negative",
                           p_value >= 0.05 ~ "not-sig"))

offset_coef_sum <- offset_coef_df %>% 
  group_by(slope) %>% 
  summarise(slope_count = n()) %>% 
  mutate(phenometric = "Termination")

# bind two datasets together
total_coef <- rbind(onset_coef_sum, offset_coef_sum)
total_coef$phenometric <- factor(total_coef$phenometric, levels = c("Onset", "Termination"))
print(total_coef)
```

## Plot results
Note results of our analysis that was generated independently show the same pattern as demonstrated in Larsen and Shirey (2021), Figure 2c & Figure 2f

```{r plotResults, echo = TRUE}
  ggplot() +
  geom_bar(total_coef,mapping = aes(x = phenometric, y = slope_count,
                                     fill = slope), stat = "identity") +
  scale_fill_manual(values = c("blue", "grey", "dark green"))
```

